---
title: "ST502 - Project 1 - Fun with GIFs!"
author: "Brian Sugg"
date: "February 2021"
output:
  html_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(formatR)
library(dplyr)
library(ggplot2)
library(gganimate)
```

# Project 1 Recap  

## Goal  

This project involves a Monte Carlo simulation study using data generated in R to investigate the properties of estimators and confidence intervals.  

The primary goal is to compare different confidence interval procedures and how successful they are in capturing *p*, the probability of success from a Binomial random sample.  

## Methods Used  

To conduct inference on *p*, this report has focused on six different methods for generating confidence intervals, including:  

* Wald interval  
* Adjusted Wald interval  
* Clopper-Pearson (exact) interval  
* Score interval  
* Raw percentile interval using a parametric bootstrap  
* Bootstrap t interval using a parametric bootstrap  

## Simulations  

Which intervals should we use in practice? We really need to understand what *good properties* of a confidence interval are, such as:  

* Proportion of intervals that capture the true value (hopefully 1-$\alpha$)  
* Average length of the interval  

By focusing on these two properties we can compare the performance of these intervals for making inference from various combinations of *p* and *n* using a simulation with `rbinom()` to generate data for our comparison.  

### General Procedure  

1. Generate *N* = 1,500 random samples from a binomial where *n* varies from 1 to 200 and *p* varies from 0.01 to 0.99 in increments of 0.01. For example, `rbinom(1500, size = 15, prob = 0.01)` does the trick for one combination of *n* = 15 and *p* = 0.01!  
2. Create and save 95% CIs for *p* using the six methods mentioned above. For the bootstrap methods, use *B* = 200 resamples.  
3. Calculate the two properties discussed previously for all combinations of n and p using the simulation results.  

The entire simulation procedure and resulting calculations for all combinations of *n* and *p*, where *n* = 1 to 200 and *p* varies from 0.01 to 0.99 in increments of 0.01, takes around 10 hours (19,800 possible combinations).  

# GIF Animation

A GIF (Graphical Interchange Format) is a small collection of images that can serve as a powerful tool for bringing data to life, especially when animating a time series or other iterative type of data set.  

When paired together, the `ggplot2` and `gganimate` packages offer a simple method for transforming static plots into animated GIFs!  

```{r packages}
library(ggplot2)
library(gganimate)
```

## Data Load

The simulation results from Project 1 containing all 19,800 possible combinations of *n* and *p* have been saved in a `.RData` file. Begin the GIF creation process by saving this file locally, and then loading it into your R session using code similar to the below.  

```{r loadData}
# Set working directory and load stored data generated from the simulation
setwd("~/ST502/Projects/Project1/gif")
load("binom95CI.RData")
```

The `binom95CI.RData` file contains three data sets:  

* `dataSet` - table of results for all 19,800 combinations of *n* and *p*  
* `plotContain` - table of results in plot-friendly format for proportion of intervals that capture the true value of *p*  
* `plotLength` - table of results in plot-friendly format for the average length of intervals  

## Plot Creation

For context, a typical static plot can be created with the `ggplot2` package. An example has been provided below using subsets of `plotContain` and `plotLength` where *n* = 15.

```{r createContainPlot}
# Create plot for Proportion Contained
## Optional: Use data=subset(plotContain,Method=="") to plot specific methods
lineContain <- ggplot(data=subset(plotContain,n==15),aes(x=p,y=value,color=Method))
lineContain <- lineContain +
  geom_line(size=1) +
  scale_y_continuous(breaks = c(0,.2,.4,.6,.8,.95), limits = c(0,1)) +
  facet_wrap(vars(Method),labeller = label_both) +
  labs(x="p",y="Proportion Contained") +
  theme_bw() +
  ggtitle("Proportion Contained @ n: 15")
lineContain
```

```{r createLengthPlot}
# Create plot for Average Interval Length
## Optional: Use data=subset(plotLength,Method=="") to plot specific methods
lineLength <- ggplot(data=subset(plotLength,n==15),aes(x=p,y=value,color=Method))
lineLength <- lineLength +
  geom_line(size=1) +
  scale_y_continuous(breaks = c(0,.2,.4,.6,.8,1), limits = c(0,1)) +
  labs(x="p",y="Average Interval Length") +
  theme_bw() +
  ggtitle("Average Interval Length @ n: 15")
lineLength
```

What if we wanted to study the results of the other possible values of *n* without having to manually create multiple charts? Essentially, how can we illustrate the impact from an increase in sample size, *n*?

**By generating a GIF with the help of `gganimate`!**  

## GIF Creation

By introducing a few additional lines of code when defining the attributes of our `ggplot` object, we can utilize the power of the `gganimate` package to render images into a GIF animation for all desired values of *n*.  

The `animate()` function is used for customizing various properties of the generated GIF.  

Additionally, any created GIF can be saved locally for future sharing over email, discussion boards, etc with the `anim_save()` function.  

### For *n* = 1 to 50

Rendering a GIF can sometimes take several minutes depending on the number of needed frames and individual plots that have to be created. For an example of a shorter GIF that can still convey key trends as *n* increases, try a subset of the plot data where *n* goes from 1 to 50.

```{r createContainGIF50}
# Create GIF for Proportion Contained
lineContain50 <- ggplot(data=subset(plotContain,n<=50),aes(x=p,y=value,color=Method))
lineContain50 <- lineContain50 +
  geom_line(size=1) +
  scale_y_continuous(breaks = c(0,.2,.4,.6,.8,.95), limits = c(0,1)) +
  facet_wrap(vars(Method),labeller = label_both) +
  labs(x="p",y="Proportion Contained") +
  theme_bw() +
  # gganimate code below
  ggtitle("Proportion Contained @ n: {frame_time}") +
  transition_time(n) + # Generates frame_time var used in ggtitle()
  ease_aes("linear") +
  enter_fade() + # Optional
  exit_fade() # Optional

# Animate the ggplot object
animate(lineContain50,width=960,height=540)

# Set filename and save most recent animation to current working directory
anim_save("Contain_nTo50.gif")
```

```{r createLengthGIF50}
# Create GIF for Average Interval Length
lineLength50 <- ggplot(data=subset(plotLength,n<=50),aes(x=p,y=value,color=Method))
lineLength50 <- lineLength50 +
  geom_line(size=1) +
  scale_y_continuous(breaks = c(0,.2,.4,.6,.8,1), limits = c(0,1)) +
  labs(x="p",y="Average Interval Length") +
  theme_bw() +
  # gganimate code below
  ggtitle("Average Interval Length @ n: {frame_time}") +
  transition_time(n) + # Generates frame_time var used in ggtitle()
  ease_aes("linear") +
  enter_fade() + # Optional
  exit_fade() # Optional

# Animate the ggplot object
animate(lineLength50,width=960,height=540)

# Set filename and save most recent animation to current working directory
anim_save("Length_nTo50.gif")
```

### For *n* = 1 to 200

For further analysis, the entire simulation can be illustrated as *n* goes from 1 to 200.  

**Important Notes:**  

* The `subset()` function has been removed from the `ggplot(data=)` declaration, and all values of *n* are rendered.  
* A new option, `nframes=` has been declared in the `animate()` function to force the creation of 200 frames within the animation, one for each value of *n*. Without this option, `gganimate` could create an animation that skips every other value of *n*, resulting in a less-desirable, choppy animation.  

```{r createContainGIF200}
# Create GIF for Proportion Contained
lineContain200 <- ggplot(data=plotContain,aes(x=p,y=value,color=Method))
lineContain200 <- lineContain200 +
  geom_line(size=1) +
  scale_y_continuous(breaks = c(0,.2,.4,.6,.8,.95), limits = c(0,1)) +
  facet_wrap(vars(Method),labeller = label_both) +
  labs(x="p",y="Proportion Contained") +
  theme_bw() +
  # gganimate code below
  ggtitle("Proportion Contained @ n: {frame_time}") +
  transition_time(n) + # Generates frame_time var used in ggtitle()
  ease_aes("linear") +
  enter_fade() + # Optional
  exit_fade() # Optional

# Animate the ggplot object
animate(lineContain200,width=960,height=540,nframes=200)

# Set filename and save most recent animation to current working directory
anim_save("Contain_nTo200.gif")
```

```{r createLengthGIF200}
# Create GIF for Average Interval Length
lineLength200 <- ggplot(data=plotLength,aes(x=p,y=value,color=Method))
lineLength200 <- lineLength200 +
  geom_line(size=1) +
  scale_y_continuous(breaks = c(0,.2,.4,.6,.8,1), limits = c(0,1)) +
  labs(x="p",y="Average Interval Length") +
  theme_bw() +
  # gganimate code below
  ggtitle("Average Interval Length @ n: {frame_time}") +
  transition_time(n) + # Generates frame_time var used in ggtitle()
  ease_aes("linear") +
  enter_fade() + # Optional
  exit_fade() # Optional

# Animate the ggplot object
animate(lineLength200,width=960,height=540,nframes=200)

# Set filename and save most recent animation to current working directory
anim_save("Length_nTo200.gif")
```

What trends and observations are more apparent from the animations compared to the static plot images? Enjoy!  
